### 2021-06-08 thesis/tex/ICintro.tex
update

In this work, we only study the detectors handeled by the IceCube collaboration. They are of Cherenkov type, which means that they detect 
the secondary charged lepton by its emitted Cherenkov light, produced from its travel through the Antarctic ice.

If the charged leptons interact heavily with the ice, they will travel a short distance and emit a localized flash of 
Cherenkov light. This event is referred to as a cascade. The neutral current interactions involves quarks, which recoils and procudes
showers of hadrons. Also, charged current $\ne$ interactions also produce cascades. A cascade event 
is shown in Fig.~\ref{fig:events_cascade}.

If the charged leptons don't interact as much in the ice, they penetrate a larger part of it, emitting light and tertiary particles
as they go. This event is referred to as a track, and are often due to muon charged current interactions. A track event 
is shown in Fig.~\ref{fig:events_track}.

@@

\subsection{Atmospheric neutrino flux}
Atmospheric neutrinos originates from cosmic rays composed of protons interacting with nuclei in the atmosphere.
These interactions ultimately produces pions, which decay as 
\begin{align}\label{eq:pion}
    &\pi^+ \to \mu^+ + \nm\,, \quad \pi^- \to \mu^- + \anm \nonumber \\
    &\pi^+ \to e^+ + \ne\,,\,\, \quad \pi^- \to e^- + \ane\,.
\end{align}
In the muonic decay channel, muons are emitted which will be detected by the IceCube detector. A part of the 
uncorrelated systematic error comes from this \emph{muon background}, i.e. events misclassified as muons from 
$\nm$ interactions rather than from pion decay. Moreover, the atmospheric flux is often associated with a large error.
In this work, we will use a flux normalization error of 24\%, and a zenith slope error of 4\%~\cite{hondapaper}.

The flux is provided in \cite{hondaData,hondaArticle}, and a selection is shown in Table~\ref{table:flux}
The flux data is binned in $\ztrue$. The fluxes are averaged over azimuthal direction and over solar minimum/maximum. 
The units of the fluxes are given as \si{\per\GeV \per\metre\squared \per\second \per\steradian} and are omitted
from the table for clarity. 
We note that the fluxes for $\nt$ and $\nu_{\bar{\tau}}$ are missing. Kaons can decay into neutral pions, which in turn 
can produce $\nt$, but this branching ratio is extremely small. Thus, we never have to use probabilities on the form 
$P_{\tau \beta}$, since we have no incoming atmospheric $\nu_\tau$ flux.

Interpolating the data yields makes us capable of returning all four necessary fluxes for a given true energy and true zenith.
The result is shown in Fig.~\ref{fig:flux_aeff}.

@@

\subsection{Event reconstruction}
After an event has occurred, the IceCube algorithms process the data coming from the detector to \emph{reconstruct} the event. This means that, given the parameters recorded by the detector, what are their "true" values?
We are interested in two variables: the energy and the direction. Each event is tagged with a probable energy and zenith angle, called the recostructed parameters $\Ereco$ and $\zreco$, which are the parameters according to the DOMs.
The collaboration then uses numerous sophisticated methods to backtrack the reconstructed parameters to the true parameters. So a charged lepton hits the DOMs, and we ultimately end up with the associated neutrino's true and reconstructed energy and zenith angle. The reconstructed parameters are what we are using to analyze the data (because this is what the detector actually sees), while the true parameters are used in the determination of that neutrino's "actual" flux and cross-section (because this is what nature sees).

>>>>

In this work, we only study the detectors handeled by the IceCube collaboration. They are of Cherenkov type, which means that they detect 
the secondary charged lepton by its emitted Cherenkov light, produced from its travel through the Antarctic ice. 
If the charged leptons interact heavily with the ice, they will travel a short distance and emit a localized flash of 
Cherenkov light. This event is referred to as a cascade. The neutral current interactions involves quarks, which recoils and procudes
showers of hadrons. Also, charged current $\ne$ interactions also produce cascades. A cascade event 
is shown in Fig.~\ref{fig:events_cascade}.
If the charged leptons don't interact as much in the ice, they penetrate a larger part of it, emitting light and tertiary particles
as they go. This event is referred to as a track, and are often due to muon charged current interactions. A track event 
is shown in Fig.~\ref{fig:events_track}.

@@

\subsection{Atmospheric Neutrino Flux}
Atmospheric neutrinos originates from cosmic rays composed of protons interacting with nuclei in the atmosphere.
These interactions ultimately produces pions, which decay as 
\begin{align}\label{eq:pion}
    &\pi^+ \to \mu^+ + \nm\,, \quad \pi^- \to \mu^- + \anm \nonumber \\
    &\pi^+ \to e^+ + \ne\,,\,\, \quad \pi^- \to e^- + \ane\,.
\end{align}
In the muonic decay channel, muons are emitted which will be detected by the IceCube detector. A part of the 
uncorrelated systematic error comes from this \emph{muon background}, i.e. events misclassified as muons from 
$\nm$ interactions rather than from pion decay. Moreover, the atmospheric flux is often associated with a large error.
In this work, we will use a flux normalization error of 24\%, and a zenith slope error of 4\%~\cite{hondapaper}.

The flux is provided in \cite{hondaData,hondaArticle}, and a selection is shown in Table~\ref{table:flux}
The flux data is binned in $\ztrue$. The fluxes are averaged over azimuthal direction and over solar minimum/maximum. 
The units of the fluxes are given as \si{\per\GeV \per\metre\squared \per\second \per\steradian} and are omitted
from the table for clarity. 
We note that the fluxes for $\nt$ and $\nu_{\bar{\tau}}$ are missing. Kaons can decay into neutral pions, which in turn 
can produce $\nt$, but this branching ratio is extremely small. Thus, we never have to use probabilities on the form 
$P_{\tau \beta}$, since we have no incoming atmospheric $\nu_\tau$ flux. 
Interpolating the data yields makes us capable of returning all four necessary fluxes for a given true energy and true zenith.
The result is shown in Fig.~\ref{fig:flux_aeff}.

@@

\subsection{Event Reconstruction}
After an event has occurred, the IceCube algorithms process the data coming from the detector to \emph{reconstruct} the event. This means that, given the parameters recorded by the detector, what are their "true" values?
We are interested in two variables: the energy and the direction. Each event is tagged with a probable energy and zenith angle, called the recostructed parameters $\Ereco$ and $\zreco$, which are the parameters according to the DOMs.
The collaboration then uses numerous sophisticated methods to backtrack the reconstructed parameters to the true parameters. So a charged lepton hits the DOMs, and we ultimately end up with the associated neutrino's true and reconstructed energy and zenith angle. The reconstructed parameters are what we are using to analyze the data (because this is what the detector actually sees), while the true parameters are used in the 
determination of that neutrino's `actual' flux and cross-section (because this is what Nature sees).

### 2021-06-08 thesis/tex/SM.tex
update

% \input{commands}\begin{document}
\section{The Standard Model}\label{ch:SM}
In order to describe the three forces of nature, we gather the mediators of each force -- the vector bosons -- into gauge groups. 
Each vector boson has one corresponding generator, the set of which constitutes the group.
The strong charge is mediated by eight massless gluons, which correspond to the eight generators of $\text{SU}(3)_C$. 
The weak charge is mediated by the three massive gauge bosons $W^\pm$ and $Z$ and the massless boson $\gamma$, 
which constitute the generators of $\text{SU}(2)_L$ and $\text{U}(1)_Y$.

The subscript of each group denotes by which mechanism that force is mediated. The gluons mediate the strong force through interactions of color, emphasized with subscript $C$. The weak force only sees left-handed particles, 
which we distinguish with the subscript $L$. And the electroweak interaction that a particle undergoes is determined by its hypercharge, $Y$. For example, the quarks all have a nonzero color and hypercharge, 
so they participate in the strong and electromagnetic interactions. If a quark is left-handed, it will also feel the weak interaction. The neutrinos on the other hand have neither charge nor color, so they are insivible to both the strong and electromagnetic force.

Together, these groups make up the local (gauge) symmetry group $\mathrm{SU}(3)_{\mathrm{C}} \times \mathrm{SU}(2)_{L} \times \mathrm{U}(1)_{Y}$. 
This symmetry group determines the form of the three coupling constants of nature, whose numerical values must be experimentally measured. 
Since the vector bosons are represented by the generators, they are wholly constrained by the symmetry group. However, the scalar bosons and fermions are free as long as they
belong to representations of the symmetry group. By this construction, the fermions have more degrees of freedom with which we can propose amendments to the model with. Even the
number of fermions must be experimentally derived. In this work,
we will use this leniency to make drastic modifications to the Standard Model fermions and examine to which extent these theories might be falsifiable.

\subsection{Lepton mixing}
Now we shall see how the charged leptons interact with the chargeless neutrinos.
Consider how charged leptons interact with the Higgs field. In the unitary gauge, the Higgs-lepton Yukawa Lagrangian 
\begin{align}
    \label{eq:YukawaLagrangian}
    \mathcal{L}_{H}=-\left( \frac{v + H}{\sqrt{2}} \right) \ell_{\alpha L}^{\prime} Y_{\alpha \beta}^{\prime \ell} \ell_{\beta R}^{\prime}
\end{align}
has a non-diagonal Yukawa coupling matrix $Y^{\prime \ell}$, which results in the charged lepton masses being non-definite. This can be remedied by diagonalizing $Y^{\prime \ell}$ with a unitary matrix $V^\ell$
\begin{align}\label{eq:leptonYukawaDiag}
    V_{L}^{\ell \dagger} Y^{\prime \ell} V_{R}^{\ell}=Y^{\ell}\,.
\end{align}
This procedure give the charged lepton fields $\ell_{X}=V_{X}^{\ell \dagger} \ell_{X}^{\prime}$ definite mass. 
However, this alters the form of the neutral and charged current interactions, which the neutrinos undergo. The leptonic charged weak current now takes the form
\begin{align}
    \label{eq:j_CC}
    j^\rho_W &= 2 \bar{\nu}^\prime_\alpha \gamma^\rho \ell^\prime_\alpha \nonumber \\
             &= 2 \bar{\nu}_\alpha V^{\ell \dagger} \gamma^\rho V^\ell \ell_\alpha \nonumber \\
             &= 2 \bar{\nu}_\alpha \gamma^\rho \ell_\alpha\,,
\end{align}
where each lepton field and matrix have an implied $L$ subscript, which was dropped for clarity. The flavor neutrino fields $\nu_\alpha$ now only couple to the associated charged lepton field $\ell_\alpha$. However, 
$\nu_\alpha$ remained massless, since no flavor rotation $V^{\ell \dagger}$ can make massless fields massive.
%TODO:more here about SM and lepton mixing? maybe about mass/massless particles. why does SM predict massless neutrinos, and 
%what implications does it have?
% \bibliographystyle{apalike}
% \bibliography{ref}
% \end{document}

>>>>

% \input{commands}\begin{document}
\begin{figure}[!h]
    \centering
    \includegraphics[width=0.7\textwidth]{figures/SM_figure.pdf}
\end{figure}
\section{The Standard Model}\label{ch:SM}
In order to describe the three quantizable forces of nature, we gather the mediators of each force -- the vector bosons -- into local (gauge) symmetry groups. 
Each vector boson has one corresponding generator, the set of which constitutes the group.
The strong charge is mediated by eight massless gluons, which correspond to the eight generators of $\text{SU}(3)_C$. 
The weak charge is mediated by the three massive gauge bosons $W^\pm$ and $Z$, and the massless photon $\gamma$, 
which constitute the generators of $\text{SU}(2)_L$ and $\text{U}(1)_Y$.

The subscript of each group denotes by which mechanism that force is mediated. The gluons mediate the strong force through interactions of color, emphasized with subscript $C$. The weak force only sees left-handed particles, 
which we distinguish with the subscript $L$. And the electroweak interaction that a particle undergoes is determined by its hypercharge $Y$. For example, the quarks all have a nonzero color and hypercharge, 
so they participate in the strong and electromagnetic interactions. If a quark is left-handed, it will also feel the weak interaction. The neutrinos on the other hand, have neither charge nor color, 
so they are insivible to both the strong and electromagnetic force. We express this by letting their fields transform as singlets under those symmetry groups.

Together, these three interactions make up the Standard Model gauge group $\mathrm{SU}(3)_{\mathrm{C}} \times \mathrm{SU}(2)_{L} \times \mathrm{U}(1)_{Y}$. 
This determines the form of the three coupling constants, but which numerical values must be experimentally measured. 
Since the vector bosons are represented by the generators, they are uniquely determined by the symmetry group. However, the scalar boson(s) and fermions are free as long as they
belong to representations of the symmetry group. By this construction, the fermions have more degrees of freedom with which we can propose amendments to the model with. Even the
number of fermions must be experimentally verified and can be altered from a phenomenological standpoint. In this work,
we will use this leniency to make drastic modifications to the Standard Model fermions and examine to which extent 
these modifications might be supported by experimental evidence.

% \subsection{Lepton mixing}
% Now we shall see how the charged leptons interact with the chargeless neutrinos.
% Consider how charged leptons interact with the Higgs field. In the unitary gauge, the Higgs-lepton Yukawa Lagrangian 
% \begin{align}
%     \label{eq:YukawaLagrangian}
%     \mathcal{L}_{H}=-\left( \frac{v + H}{\sqrt{2}} \right) \ell_{\alpha L}^{\prime} Y_{\alpha \beta}^{\prime \ell} \ell_{\beta R}^{\prime}
% \end{align}
% has a non-diagonal Yukawa coupling matrix $Y^{\prime \ell}$, which results in the charged lepton masses being non-definite. This can be remedied by diagonalizing $Y^{\prime \ell}$ with a unitary matrix $V^\ell$
% \begin{align}\label{eq:leptonYukawaDiag}
%     V_{L}^{\ell \dagger} Y^{\prime \ell} V_{R}^{\ell}=Y^{\ell}\,.
% \end{align}
% This procedure give the charged lepton fields $\ell_{X}=V_{X}^{\ell \dagger} \ell_{X}^{\prime}$ definite mass. 
% However, this alters the form of the neutral and charged current interactions, which the neutrinos undergo. The leptonic charged weak current now takes the form
% \begin{align}
%     \label{eq:j_CC}
%     j^\rho_W &= 2 \bar{\nu}^\prime_\alpha \gamma^\rho \ell^\prime_\alpha \nonumber \\
%              &= 2 \bar{\nu}_\alpha V^{\ell \dagger} \gamma^\rho V^\ell \ell_\alpha \nonumber \\
%              &= 2 \bar{\nu}_\alpha \gamma^\rho \ell_\alpha\,,
% \end{align}
% where each lepton field and matrix have an implied $L$ subscript, which was dropped for clarity. The flavor neutrino fields $\nu_\alpha$ now only couple to the associated charged lepton field $\ell_\alpha$. However, 
% $\nu_\alpha$ remained massless, since no flavor rotation $V^{\ell \dagger}$ can make massless fields massive.
% \bibliographystyle{apalike}
% \bibliography{ref}
% \end{document}

### 2021-06-08 thesis/tex/nsiTheory.tex
update

% \begin{document}
\section{Non-Standard Interactions}
We have already seen how the failure of the Standard Model to predict neutrino masses and flavor 
osillations forces us to amend it. We have no reason to believe that the Standard Model 
gauge group is the complete picture. Just as the electroweak theory $SU(2)_L \times U(1)_Y$ is 
spontaneously broken to $U(1)_{EM}$, a higher order theory at a different energy scale with ompletely different
properties might be responsible for the Standard Model that we know. In this sense, the Standard Model might be considered 
an effective low energy theory, which is yiels impressive results in some areas, but failing in others. In fact, 
just as the fermion masses originate from the electroweak symmetry breaking, the neutrino masses might be generated from 
another broken symmetry resulting in the Standard Model gauge group.

%TODO: moreintro. incompleteness of SM
Following the discussion in~\cite{tommyNSI}, the new parameters $\epsilon$ have the scale
\begin{align}
    \epsilon \propto \frac{m_W^2}{m_{\epsilon}^2} \sim \frac{10^{-2}}{m_\epsilon^2}\,
\end{align} 
in TeV, so the new interactions generated at a mass scale of $m_\epsilon = \SI{1}{TeV}$ will produce parameters in the order of $10^{-2}$, 
two magnitudes below the standard matter effect. Thus, if we assume the new interactions to arise from a higher-energy theory at or above electroweak scale, 
we then predict that the parameters contribute with at most a factor of $10^{-2}$ to the standard matter effect, decreasing quadratically.

\subsection{Altering the Matter Potential}
Up until now, we have only considered weak neutrino interactions with electrons, protons, and neutrons.
Following the approach of the Standard Model that the group generators uniquely determine the gauge bosons, a different gauge
theory will have different interactions than those we presently know. We call these interactions non-standard interactions (NSI).
We can phenomenologically allow these interactions to include the up and down quarks which are present 
in the Earth as the fundamental components of neutrons and protons, 
parametrized as model-independent four-fermion effective operators \cite{salvadoNSI,nsiFarzan}.
We write the Lagrangians
\begin{align*}
    \mathcal{L}_{\mathrm{CC}} &= -2 \sqrt{2} G_{F} \epsilon_{\alpha \beta}^{f f^{\prime} X}\left(\bar{\nu}_{\alpha} \gamma^{\mu} P_{L} \ell_{\beta}\right)\left(\bar{f}^{\prime} \gamma_{\mu} P_{X} f\right) \\
    \mathcal{L}_{\mathrm{NC}} &= -2 \sqrt{2} G_{F} \epsilon_{\alpha \beta}^{f X}\left(\bar{\nu}_{\alpha} \gamma^{\mu} P_{L} \nu_{\beta}\right)\left(\bar{f} \gamma_{\mu} P_{X} f\right)\,,
\end{align*}
where CC denotes the charged current interaction with the matter field $f\neq f^\prime \in \{u,d\}$, and NC denotes the neutral current interaction with 
$f \in \{e,u,d\}$. The CC NSI effect affects neutrino production and detection, and will not be considered here. NC NSI affect the matter potential, 
and is thus of interest to us.

We have no independent sensitivity for the neither chirality nor flavor type of $\epsilon^X$, so we sum over these and study the effective matter NSI parameter
 $\epsilon_{\ab}$:
\begin{align}
    \epsilon_{\ab} = \sum_{X \in \{L,R\}} \sum_{f \in \{e,u,d\}} \frac{N_f}{N_e} \epsilon^{fX}_{\ab}\,.
\end{align}
Our matter study will be wholly confined to the interior of the Earth, where we assume electrical neutrality and equal distribution of neutrons and protons, 
we get $N_u/N_e \simeq N_d/N_e \simeq 3$. Also we assume the components $\epsilon_{\a\b}$ to be real. Thus,
\begin{align}
    \epsilon_{\ab} =  \sum_X \epsilon_{\ab}^{eX} + 3(\epsilon_{\ab}^{uX} + \epsilon_{\ab}^{dX})
\end{align}
Now, $\epsilon_{\ab}$ enters the Hamiltonian as entries of a potential-like matrix. In Eq.~\ref{eq:NSIH}, $A_{CC}\text{diag}(1,0,0)$ is our 
familiar matter potential from the Standard Model. There is also our new term, $A_{CC} \epsilon$, which contains the components $\epsilon_{\a\b}$:
\begin{align}\label{eq:NSIH}
    H &= \frac{1}{2E} \left[UM^2U^\dagger + A_{CC}\text{diag}(1,0,0) + A_{CC} \epsilon \right] \nonumber \\
      &= \frac{1}{2E} \left[UM^2U^\dagger + A_{CC}
      \begin{pmatrix}
          1 + \epsilon_{ee} & \epsilon_{e\mu} & \epsilon_{e\tau}  \\
          \epsilon_{\mu e} & \epsilon_{\mu\mu} & \epsilon_{\mu\tau}  \\
          \epsilon_{\tau e} & \epsilon_{\tau\mu} & \epsilon_{\tau\tau}
      \end{pmatrix} \right]\,.
\end{align} 
In the limit $\epsilon_{\ab} \to 0$, we recover the standard interaction Hamiltonian from Eq.~\ref{eq:H_3gen}.
We can draw several conclusions from this form of the Hamiltonian. Any nonzero off-diagonal element $\epsilon_{\ab}, \a \neq \b$ will make the NSI violate 
lepton flavor, just as the off-diagonal elements of $U$ does in the SM. Moreover, since the SM potential has the same order in $A_{CC}$ as the NSIs, 
any $\epsilon_{\ab} \sim 1$ will make the new matter effect be the same order as the SM effect.

@@

\subsection{NSIs in IceCube}
In our analysis of IceCube, we are constrained to muon track events. Thus, we are not able to test any theory which does not modify $P_{\a \mu}$. Moreover,
the IceCube data is avaibable in the range \SI{500}{\GeV} to \SI{10}{\TeV} range, where any rapid oscillations have settled down.

@@

Moving on to $\eem$ and Fig.~\ref{fig:eem_eet_probs}, we see that both probabilities has shifted downwards for $\Etrue > \SI{500}{\GeV}$.
In Fig.~\ref{fig:eem_eet_probs}, we see that the muon channel remains largely unaffected of the value of $\eet$ as we expected. The expection of this lies 
in the DeepCore region of rapid oscillations, where mixing is more violent. 
\begin{figure}\label{fig:emt_ett_probs}
    \begin{center}%TODO subplots and fix labels in parahgraph aboveto match them
        \includegraphics[width=0.45\textwidth]{figures/emt_probs.pdf}
        \includegraphics[width=0.45\textwidth]{figures/ett_probs.pdf}
        \caption{\emph{Left panel:} Muon neutrino and antineutrino survival probabilities for
        $\ztrue = -1$ when $\emt = 0.02$. All other NSI parameters are fixed to zero.
        \emph{Right panel:} Muon neutrino and antineutrino survival probabilities for
        $\ztrue = -1$ when $\ett = 0.05$. All other NSI parameters are fixed to zero.}
    \end{center}
\end{figure}

\begin{figure}\label{fig:eem_eet_probs}
    \begin{center}
        \includegraphics[width=0.45\textwidth]{figures/eem_probs.pdf}
        \includegraphics[width=0.45\textwidth]{figures/eet_probs.pdf}
        \caption{Muon neutrino and antineutrino survival probabilities for
        $\ztrue = -1$\emph{Left panel:} $\eem = 0.2$. All other NSI parameters are fixed to zero.
        \emph{Right panel:} $\eet = 0.2$. All other NSI parameters are fixed to zero.}
    \end{center}
\end{figure}

\subsection{NSIs in DeepCore and PINGU}
Now we repeat our proability analysis but for the DeepCore/PINGU region of \SIrange{5.6}{56}{\GeV}. As we previously saw,
we have rapid oscillations, which means that "indirect" modifications (i.e. $\eet$ will affect the $\Pmm$ channel)
will be more apparent, since all flavors are involved to a greater degree compared with the more stable region above \SI{500}{\GeV}, where 
many oscillations have settled down.

@@

We plot the event pull $(N_{NSI} - N_{SI})/\sqrt{N_{SI}}$ where $N_{(N)SI}$ are the numbers of expected events
assuming (non-)standard interactions in Fig.~\ref{fig:event_pulls}. This gives the normalized difference in the
number of expected events at the detector, and illustrates the expected sensitivity of DeepCore for the NSI parameters.
%TODO: clean up above and expand a bit.
\begin{figure}\label{fig:Pee_eet_probs}
    \begin{center}
        \includegraphics[width=0.6\textwidth]{figures/Pee_eet_probs.pdf}
        \caption{Electron neutrino and antineutrino survival probabilities for
        $\ztrue = -1$ when $\eet = 0.2$. All other NSI parameters are fixed to zero.}
    \end{center}
\end{figure}

>>>>

% \begin{document}
\section{Non-Standard Interactions}\label{sec:nsiTheory}
We have no reason to believe that the Standard Model 
gauge group is the complete picture. We have already seen how the failure of the Standard Model to predict neutrino masses and flavor 
osillations forces us to amend it. 
Just as the electroweak theory $SU(2)_L \times U(1)_Y$ is 
spontaneously broken to $U(1)_{EM}$, a higher order theory at a different energy scale with completely different
properties might undergo a similar spontaneous symmetry break at high energies, 
producing at lower energies the Standard Model that we know. In fact, 
just as the fermion masses originate from the electroweak symmetry breaking, the neutrino masses might be generated from 
another broken symmetry resulting in the Standard Model gauge group.
In this sense, the Standard Model might be considered 
an effective low energy theory, which is yiels impressive results in some areas, but failing in others.

Compared to other fermions, the lightness of neutrino masses along with their sparse SM interactions might 
indicate that these particles provide the best starting point for us to probe new physics, in which 
other neutrino interactions might occur~\cite{gavela2009}. We call these interactions non-standard interactions (NSI).

Following the approach of the Standard Model that the group generators uniquely determine the gauge bosons, a different gauge
theory will have different interactions than those we presently know. These new interactions can be parametrized as model-independent four-fermion effective operators \cite{salvadoNSI,nsiFarzan}.
Following the discussion in~\cite{tommyNSI}, the six-dimensional NSI parameters $\epsilon$ have the scale
\begin{align}
    \epsilon \propto \frac{m_W^2}{m_{\epsilon}^2} \sim \frac{10^{-2}}{m_\epsilon^2}\,
\end{align} 
in TeV, so the new interactions generated at a mass scale of $m_\epsilon = \SI{1}{TeV}$ will produce parameters in the order of $10^{-2}$, 
two magnitudes below the standard matter effect. Thus, if we assume the new interactions to arise from a higher-energy theory above electroweak scale, 
we then predict that the parameters contribute with at most a factor of $10^{-2}$ to the standard matter effect, decreasing quadratically.

\subsection{Altering the Matter Potential}
Up until now, we have only considered weak neutrino interactions with electrons, protons, and neutrons.
We can phenomenologically allow these interactions to include the up and down quarks which are present 
in the Earth as the fundamental components of neutrons and protons, as seen in the Lagrangians
\begin{align*}
    \mathcal{L}_{\mathrm{CC}} &= -2 \sqrt{2} G_{F} \epsilon_{\alpha \beta}^{f f^{\prime} X}\left(\bar{\nu}_{\alpha} \gamma^{\mu} P_{L} \ell_{\beta}\right)\left(\bar{f}^{\prime} \gamma_{\mu} P_{X} f\right) \\
    \mathcal{L}_{\mathrm{NC}} &= -2 \sqrt{2} G_{F} \epsilon_{\alpha \beta}^{f X}\left(\bar{\nu}_{\alpha} \gamma^{\mu} P_{L} \nu_{\beta}\right)\left(\bar{f} \gamma_{\mu} P_{X} f\right)\,,
\end{align*}
where CC denotes the charged current interaction with the matter field $f\neq f^\prime \in \{u,d\}$, and NC denotes the neutral current interaction with 
$f \in \{e,u,d\}$. The CC NSI effect affects neutrino production and detection, and will not be considered further. NC NSI affect the matter potential, 
and is thus of interest to us.

We have no independent sensitivity for the neither chirality nor flavor type of $\epsilon^X$, so we sum over these and study the effective matter NSI parameter
 $\epsilon_{\ab}$:
\begin{align}
    \epsilon_{\ab} = \sum_{X \in \{L,R\}} \sum_{f \in \{e,u,d\}} \frac{N_f}{N_e} \epsilon^{fX}_{\ab}\,.
\end{align}
Our matter study will be wholly confined to the interior of the Earth, where we assume electrical neutrality and equal distribution of neutrons and protons, 
we get $N_u/N_e \simeq N_d/N_e \simeq 3$. Also we assume the components $\epsilon_{\a\b}$ to be real. Thus,
\begin{align}
    \epsilon_{\ab} =  \sum_X \epsilon_{\ab}^{eX} + 3(\epsilon_{\ab}^{uX} + \epsilon_{\ab}^{dX})
\end{align}
Now, $\epsilon_{\ab}$ enters the Hamiltonian as entries of a potential-like matrix. In Eq.~\ref{eq:NSIH}, $A_{CC}\text{diag}(1,0,0)$ is our 
familiar matter potential from the Standard Model. There is also our new term, $A_{CC} \epsilon$, which contains the components $\epsilon_{\a\b}$:
\begin{align}\label{eq:NSIH}
    H &= \frac{1}{2E} \left[UM^2U^\dagger + A_{CC}\,\text{diag}(1,0,0) + A_{CC}\, \epsilon \right] \nonumber \\
      &= \frac{1}{2E} \left[UM^2U^\dagger + A_{CC}
      \begin{pmatrix}
          1 + \epsilon_{ee} & \epsilon_{e\mu} & \epsilon_{e\tau}  \\
          \epsilon_{\mu e} & \epsilon_{\mu\mu} & \epsilon_{\mu\tau}  \\
          \epsilon_{\tau e} & \epsilon_{\tau\mu} & \epsilon_{\tau\tau}
      \end{pmatrix} \right]\,.
\end{align} 
In the limit $\epsilon_{\ab} \to 0$, we recover the standard interaction Hamiltonian from Eq.~\ref{eq:H_3gen}.
We can draw several conclusions from this form of the Hamiltonian. Any nonzero off-diagonal element $\epsilon_{\ab}, \a \neq \b$ will make the NSI violate 
lepton flavor, just as the off-diagonal elements of $U$ does in the SM. Moreover, since the SM potential has the same order in $A_{CC}$ as the NSIs, 
any $\epsilon_{\ab} \sim 1$ will make the new matter effect be the same order as the SM effect.

@@

\subsection{IceCube Signal}
In our analysis of IceCube, we are constrained to muon track events. Thus, we are not able to test any theory which does not modify $P_{\a \mu}$. Moreover,
the IceCube data is avaibable in the range \SI{500}{\GeV} to \SI{10}{\TeV} range, where any rapid oscillations have settled down.

@@

Moving on to $\eem$ and Fig.~\ref{fig:eem_eet_probs}, we see that both probabilities has shifted downwards for $\Etrue > \SI{500}{\GeV}$.
In Fig.~\ref{fig:eem_eet_probs}, we see that the muon channel remains largely unaffected of the value of $\eet$ as we expected. The expection of this lies 
in the DeepCore region of rapid oscillations, where mixing is more violent. 
\begin{figure}
    \begin{center}%TODO subplots and fix labels in parahgraph aboveto match them
        \includegraphics[width=0.49\textwidth]{figures/emt_probs.pdf}
        \includegraphics[width=0.49\textwidth]{figures/ett_probs.pdf}
        \caption{\emph{Left panel:} Muon neutrino and antineutrino survival probabilities for
        $\ztrue = -1$ when $\emt = 0.02$. All other NSI parameters are fixed to zero.
        \emph{Right panel:} Muon neutrino and antineutrino survival probabilities for
        $\ztrue = -1$ when $\ett = 0.05$. All other NSI parameters are fixed to zero.}
        \label{fig:emt_ett_probs}
    \end{center}
\end{figure}

\begin{figure}
    \begin{center}
        \includegraphics[width=0.49\textwidth]{figures/eem_probs.pdf}
        \includegraphics[width=0.49\textwidth]{figures/eet_probs.pdf}
        \caption{Muon neutrino and antineutrino survival probabilities for
        $\ztrue = -1$\emph{Left panel:} $\eem = 0.2$. All other NSI parameters are fixed to zero.
        \emph{Right panel:} $\eet = 0.2$. All other NSI parameters are fixed to zero.}
        \label{fig:eem_eet_probs}
    \end{center}
\end{figure}

\subsection{DeepCore and PINGU Signals}
Now we repeat our proability analysis but for the DeepCore/PINGU region of \SIrange{5.6}{56}{\GeV}. As we previously saw,
we have rapid oscillations, which means that "indirect" modifications (i.e. $\eet$ will affect the $\Pmm$ channel)
will be more apparent, since all flavors are involved to a greater degree compared with the more stable region above \SI{500}{\GeV}, where 
many oscillations have settled down.

@@

We plot the event pull $(N_{NSI} - N_{SI})/\sqrt{N_{SI}}$ where $N_{(N)SI}$ are the numbers of expected events
assuming (non-)standard interactions in Fig.~\ref{fig:event_pulls}. This gives the normalized difference in the
number of expected events at the detector, and illustrates the expected sensitivity of DeepCore for the NSI parameters.
%TODO: clean up above and expand a bit.
\begin{figure}
    \begin{center}
        \includegraphics[width=0.8\textwidth]{figures/Pee_eet_probs.pdf}
        \caption{Electron neutrino and antineutrino survival probabilities for
        $\ztrue = -1$ when $\eet = 0.2$. All other NSI parameters are fixed to zero.}
        \label{fig:Pee_eet_probs}
    \end{center}
\end{figure}

### 2021-06-08 thesis/tex/oscillations.tex
update

We consider a right-handed neutrino field, $\nu_R$. Since the electroweak gauge group $\text{SU}(2)_L \times U(1)_Y$ only couple to 
left-handed particles and right-handed antiparticles, $\nu_R$ transforms as a singlet under the SM symmetry 
group $\mathrm{SU}(3)_{\mathrm{C}} \times \mathrm{SU}(2)_{L} \times \mathrm{U}(1)_{Y}$. 
This neutrino field is \emph{sterile} since it doesn't participate any of the SM interactions.

>>>>

We consider a right-handed neutrino field, $\nu_R$. Since the electroweak gauge group $\text{SU}(2)_L \times U(1)_Y$ only couple to 
left-handed particles and right-handed antiparticles, $\nu_R$ transforms as a singlet under the Standard Model symmetry 
group $\mathrm{SU}(3)_{\mathrm{C}} \times \mathrm{SU}(2)_{L} \times \mathrm{U}(1)_{Y}$. 
This neutrino field is \emph{sterile} since it doesn't participate any of the Standard Model interactions.

### 2021-06-08 thesis/tex/sterileResults.tex
update

% \begin{document}
\section{The Sterile Hypothesis}
With the null ($3\nu$) hypothesis normalized to the IceCube Monte Carlo as in Eq.~\ref{eq:Nth}, we are now in good shape to study the sterile effect on the probabilities, and how that compares to data.
First, let us see how the collected data from Fig.~\ref{fig:IC_data} deviates from the predicted $3\nu$ oscillations.
\begin{figure}
    \centering
    \includegraphics[width=0.8\textwidth]{figures/IC_rates.pdf}
    \caption{}\label{fig:IC_rates}
\end{figure}
Fig.~\ref{fig:IC_rates} shows an impressive agreement to the standard $3\nu$ oscillation picture, with the largest deviations of $~6\%$
for neutrinos close to double-digit \si{\TeV} values. In zenith, the data only deviates $\pm 1\%$. Recalling the \si{\TeV} disappearance from 
Fig.~\ref{fig:sterile_resonance}, we see that we have a similar deficiency found in the data at \SI{4}{\TeV}. Thus, we expect the best fitting 
sterile hypothesis to include a $\dm[41]$ that places the resonant disappearance in the same region. From Fig.~\ref{fig:resonance_shift}, we see
that a mass of $\dm[41] = \SI{1.5}{\eV \squared}$ achieves this.

>>>>

% \begin{document}
\section{The Sterile Hypothesis}\label{sec:sterileResults}
With the null ($3\nu$) hypothesis normalized to the IceCube Monte Carlo as in Eq.~\ref{eq:Nth}, we are now in good shape to study the sterile effect on the probabilities, and how that compares to data.
First, let us see how the collected data from Fig.~\ref{fig:IC_data} deviates from the predicted $3\nu$ oscillations.
\begin{figure}
    \centering
    \includegraphics[width=0.8\textwidth]{figures/IC_rates.pdf}
    \caption{}\label{fig:IC_rates}
\end{figure}
Fig.~\ref{fig:IC_rates} shows an impressive agreement to the standard $3\nu$ oscillation picture, with the largest deviations of $~6\%$
for neutrinos close to double-digit \si{\TeV} values. In zenith, the data only deviates $\pm 1\%$. Recalling the \si{\TeV} disappearance from 
Fig.~\ref{fig:sterile_resonance}, we see that we have a similar deficiency found in the data at \SI{4}{\TeV}. Thus, we expect the best fitting 
sterile hypothesis to include a $\dm[41]$ that places the resonant disappearance in the same region. From Fig.~\ref{fig:resonance_shift}, we see
that a mass of $\dm[41] = \SI{1.5}{\eV \squared}$ achieves this.
